---
title: "Group analysis"
output:
  pdf_document: default
  html_document: default
---

# Load packages
```{r}
pacman::p_load(readr, lme4, emmeans, yarrr, ggplot2, dplyr, sjPlot, StatisticalModels)
```

# Load data
```{r}
amp = read_csv("amp.csv")
amp_all = read_csv("amp_all.csv")
lat = read_csv("lat.csv")
lat_all = read_csv("lat_all.csv")
All_behav = read_csv("All_behav.csv")

amp$ID=as.factor(amp$ID)
amp$Deviant=as.factor(amp$Deviant)
amp$DeviantLevel=as.factor(amp$DeviantLevel)
amp$Group=as.factor(amp$Group)

amp_all$ID=as.factor(amp_all$ID)
amp_all$Deviant=as.factor(amp_all$Deviant)
amp_all$DeviantLevel=as.factor(amp_all$DeviantLevel)
amp_all$Group[amp_all$Group =="CI_ex"] = "CIx"
amp_all$Group=as.factor(amp_all$Group)


lat$ID=as.factor(lat$ID)
lat$Deviant=as.factor(lat$Deviant)
lat$DeviantLevel=as.factor(lat$DeviantLevel)
lat$Group=as.factor(lat$Group)

lat_all$ID=as.factor(lat_all$ID)
lat_all$Deviant=as.factor(lat_all$Deviant)
lat_all$DeviantLevel=as.factor(lat_all$DeviantLevel)
lat_all$Group[lat_all$Group =="CI_ex"] = "CIx"
lat_all$Group=as.factor(lat_all$Group)


All_behav$ID=as.factor(All_behav$ID)
All_behav$Deviant=as.factor(All_behav$Deviant)
All_behav$DeviantLevel=as.factor(All_behav$DeviantLevel)
All_behav$Group[All_behav$Group =="CIex"] = "CIx"
All_behav$Group=as.factor(All_behav$Group)

                        
                      
```

# Hierarchical mixed effects modeling - MMN Amplitude (T2)
```{r}
M0=lmer(Amplitude.T2~1+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M1=lmer(Amplitude.T2~Deviant+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M2=lmer(Amplitude.T2~Deviant+DeviantLevel+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M3=lmer(Amplitude.T2~Deviant+DeviantLevel+Group+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M4=lmer(Amplitude.T2~Deviant*DeviantLevel+Group+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M5=lmer(Amplitude.T2~Deviant*DeviantLevel+Deviant*Group+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M6=lmer(Amplitude.T2~Deviant*DeviantLevel+Deviant*Group+DeviantLevel*Group+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M7=lmer(Amplitude.T2~0 + Deviant*Group*DeviantLevel +(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

#comparing models
anova = anova(M0,M1,M2,M3,M4,M5,M6,M7)
anova


```

## Best model
```{r}
qqnorm(resid(M4))
plot(M4)
fixef(M7)

R2GLMER(M4)
```

## Post-hoc analysis
```{r}
emm_g <- emmeans(M4, pairwise ~ Group, adjust = "none")
emm_g[[2]]

confint(rbind(emm_g[[2]]), adjust ="none")


#Per deviant - T2
emm_r <- emmeans(M5, pairwise ~ Group)
emm_rc <- emm_r[[2]]

emm_d <- emmeans(M5, pairwise ~ Deviant)
emm_dc <- emm_d[[2]]

#Group by deviant interaction

emm_r.d <- emmeans(M5, pairwise ~ Group | Deviant)
IC_r.d <- contrast(emm_r.d[[1]], method = "pairwise")

emm_d.r <- emmeans(M5, pairwise ~ Deviant | Group)
IC_d.r <- contrast(emm_d.r[[1]], method = "pairwise")

rbind(IC_r.d[c(4:6,10:12)], adjust = "none")

confint(rbind(IC_r.d[c(4:6,10:12)],  adjust = 'none'))

```

# Hierarchical mixed effects modeling - MMN Amplitude (T1) - sanity check
```{r}
M0=lmer(Amplitude.T1~1+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M1=lmer(Amplitude.T1~Deviant+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M2=lmer(Amplitude.T1~Deviant+DeviantLevel+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M3=lmer(Amplitude.T1~Deviant+DeviantLevel+Group+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M4=lmer(Amplitude.T1~Deviant*DeviantLevel+Group+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M5=lmer(Amplitude.T1~Deviant*DeviantLevel+Deviant*Group+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M6=lmer(Amplitude.T1~Deviant*DeviantLevel+Deviant*Group+DeviantLevel*Group+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M7=lmer(Amplitude.T1~Deviant*Group*DeviantLevel+(1|ID), data=amp_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

#comparing models
anova = anova(M0,M1,M2,M3,M4,M5,M6,M7)
anova

```
```{r}
emm_r <- emmeans(M5, pairwise ~ Group)
emm_rc <- emm_r[[2]]

emm_d <- emmeans(M5, pairwise ~ Deviant)
emm_dc <- emm_d[[2]]

#Per deviant - T1
emm_r <- emmeans(M5, pairwise ~ Group)
emm_rc <- emm_r[[2]]

emm_d <- emmeans(M5, pairwise ~ Deviant)
emm_dc <- emm_d[[2]]

#Time by deviant interaction

emm_r.d <- emmeans(M5, pairwise ~ Group | Deviant)
IC_r.d <- contrast(emm_r.d[[1]], method = "pairwise")

emm_d.r <- emmeans(M5, pairwise ~ Deviant | Group)
IC_d.r <- contrast(emm_d.r[[1]], method = "pairwise")

rbind(IC_r.d[c(4:6,10:12)], adjust = "none")

confint(rbind(IC_r.d[c(4:6,10:12)],  adjust = 'none'))
```


## Best model
```{r}
qqnorm(resid(M4))
plot(M4)
R2GLMER(M4)
```

## Post-hoc analysis
```{r}
emm_g <- emmeans(M4, pairwise ~ Group, adjust = "none")
emm_g[[2]]

confint(rbind(emm_g[[2]]), adjust ="none")
```

## Plot
```{r}
ID_plot_data = read_csv2("ID_plot_data_amp.csv")

pirat = ID_plot_data
pirat$Group[pirat$Time =="CIre_T1"]=gsub("CI_re", "CIre_T1", pirat$Group[pirat$Time =="CIre_T1"])
pirat$Group[pirat$Time =="CIre_T2"]=gsub("CI_re", "CIre_T2", pirat$Group[pirat$Time =="CIre_T2"])
pirat$Group=ordered(pirat$Group, levels=c("CIre_T1", "CIre_T2", "CIex", "NH"))
pirat$Deviant=ordered(pirat$Deviant, levels=c("Intensity", "Pitch", "Timbre", "Rhythm"))

pirateplot(formula = Amp ~ Group + Deviant, data = pirat, xlab = "Group",ylab="Amplitude (µV)", ylim= c(-1.8,0.4), theme=2, pal="appletv", cex.lab = 1.1, cex.axis = 1, cex.names = 0.7, avg.line.o=1, bar.f.o = .2, bean.b.o = .7, point.o = 0.3, point.pch = 1, back.col = "white")

```

# Hierarchical mixed effects modeling - Latency (T2)
```{r}
M0=lmer(Latency_peak.T2~1+(1|ID), data=lat_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M1=lmer(Latency_peak.T2~Deviant+(1|ID), data=lat_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M2=lmer(Latency_peak.T2~Deviant+DeviantLevel+(1|ID), data=lat_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M3=lmer(Latency_peak.T2~Deviant+DeviantLevel+Group+(1|ID), data=lat_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M4=lmer(Latency_peak.T2~Deviant*DeviantLevel+Group+(1|ID), data=lat_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M5=lmer(Latency_peak.T2~Deviant*DeviantLevel+Deviant*Group+(1|ID), data=lat_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M6=lmer(Latency_peak.T2~Deviant*DeviantLevel+Deviant*Group+DeviantLevel*Group+(1|ID), data=lat_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M7=lmer(Latency_peak.T2~Deviant*DeviantLevel*Group+(1|ID), data=lat_all, REML = FALSE, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

#comparing models
anova = anova(M0,M1,M2,M3,M4,M5,M6, M7)
anova


```

## Best model
```{r}
qqnorm(resid(M5))
plot(M5)
R2GLMER(M5)
```

## Post-hoc analysis
```{r}

emm_g <- emmeans(M5, pairwise ~ Group)
emm_gc <- emm_g[[2]]

emm_d <- emmeans(M5, pairwise ~ Deviant)
emm_dc <- emm_d[[2]]

#Group by deviant interaction

emm_g.d <- emmeans(M5, pairwise ~ Group | Deviant)
IC_g.d <- contrast(emm_g.d[[1]], method = "pairwise")

rbind(IC_g.d, adjust = "none")
confint(rbind(IC_g.d), adjust ="none")

```


## Plot
```{r}
ID_plot_data = read_csv2("ID_plot_data_lat.csv")

pirat = ID_plot_data
pirat$Group[pirat$Time =="CIre_T1"]=gsub("CI_re", "CIre_T1", pirat$Group[pirat$Time =="CIre_T1"])
pirat$Group[pirat$Time =="CIre_T2"]=gsub("CI_re", "CIre_T2", pirat$Group[pirat$Time =="CIre_T2"])
pirat$Group=ordered(pirat$Group, levels=c("CIre_T1", "CIre_T2", "CIex", "NH"))
pirat$Deviant=ordered(pirat$Deviant, levels=c("Intensity", "Pitch", "Timbre", "Rhythm"))

pirateplot(formula = Lat ~ Group + Deviant, data = pirat, xlab = "Group",ylab="Latency (ms)", ylim= c(100,250), theme=2, pal="appletv", cex.lab = 1.1, cex.axis = 1, cex.names = 0.7, avg.line.o=1, bar.f.o = .2, bean.b.o = .7, point.o = 0.3, point.pch = 1, back.col = "white")


```

# Hierarchical mixed effects modeling - Behavioral (T2)
```{r}
M0=glmer(cbind(AvCor*6,6)~1+(1|ID), data=All_behav[All_behav$Group!="CIre_T1",], family="binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M1=glmer(cbind(AvCor*6,6)~Deviant+(1|ID), data=All_behav[All_behav$Group!="CIre_T1",], family="binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M2=glmer(cbind(AvCor*6,6)~Deviant+DeviantLevel+(1|ID), data=All_behav[All_behav$Group!="CIre_T1",], family="binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M3=glmer(cbind(AvCor*6,6)~Deviant+DeviantLevel+Group+(1|ID), data=All_behav[All_behav$Group!="CIre_T1",], family="binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M4=glmer(cbind(AvCor*6,6)~Deviant*DeviantLevel+Group+(1|ID), data=All_behav[All_behav$Group!="CIre_T1",], family="binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M5=glmer(cbind(AvCor*6,6)~Deviant*DeviantLevel+Deviant*Group+(1|ID), data=All_behav[All_behav$Group!="CIre_T1",], family="binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M6=glmer(cbind(AvCor*6,6)~Deviant*DeviantLevel+Deviant*Group+DeviantLevel*Group+(1|ID), data=All_behav[All_behav$Group!="CIre_T1",], family="binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

M7=glmer(cbind(AvCor*6,6)~Deviant*DeviantLevel*Group+(1|ID), data=All_behav[All_behav$Group!="CIre_T1",], family="binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))



#Comparing models 
anova = anova(M0, M1, M2, M3, M4, M5, M6, M7)
anova


#Best model
R2GLMER(M5)


```

## Post-hoc analysis
```{r}

emm_g <- emmeans(M5, pairwise ~ Group)
emm_gc <- emm_g[[2]]

emm_d <- emmeans(M5, pairwise ~ Deviant)
emm_dc <- emm_d[[2]]

#Group by deviant interaction


emm_g.d <- emmeans(M5, pairwise ~ Group | Deviant, type = "response")
IC_g.d <- contrast(emm_g.d[[1]], method = "pairwise")

rbind(IC_g.d, adjust = "none")
confint(rbind(IC_g.d), adjust ="none")

```

## Plots behavioral
```{r}
ID_plot_data = read_csv("ID_plot_data_behav.csv")

ID_plot_data$Group=ordered(ID_plot_data$Group, levels=c("CIre_T1", "CIre_T2", "CIex", "NH"))
ID_plot_data$Deviant=ordered(ID_plot_data$Deviant, levels=c("Intensity", "Pitch", "Timbre", "Rhythm"))

## Creating pirateplot
pirateplot(formula = AvCor ~ Group + Deviant, data = ID_plot_data, xlab = "Group",ylab="Average % correct", ylim= c(0,100), theme=2, pal="appletv", cex.lab = 1.1, cex.axis = 1, cex.names = 0.7, avg.line.o=1, bar.f.o = .2, bean.b.o = .7, point.o = 0.3, point.pch = 1, back.col = "white") +
  abline(a=33,33, b=0, lwd=1, lty="dashed")

```


